Empirical Bayes estimation of the normal inverse Gaussian prior in simultaneous-equation models

### 2019-11-19 docs/manuscript2.Rnw
redoing simulations

% load packages
<<packages, include=FALSE, echo=FALSE>>=
library(cambridge)
@

% set options
<<settings, include=FALSE, echo=FALSE>>=
opts_knit$set(root.dir="..", base.dir="../figs")
opts_chunk$set(fig.align='center', fig.path="../figs/", 
               echo=FALSE, cache=FALSE, message=FALSE, fig.pos='!ht')
@

@@

\documentclass[a4paper,hidelinks]{article}
\usepackage{url,bm,algorithm,algpseudocode,bbm,amsmath,amssymb,amsthm,
amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,
titlesec,multicol,verbatim,threeparttable,booktabs,dsfont,parskip,
xr-hyper,hyperref,enumerate}
\externaldocument[sm-]{supplement}

@@

% settings
\graphicspath{{/Users/magnusmunch/Documents/OneDrive/PhD/project_cambridge/graphs/}}
\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
\setlength{\textwidth}{15cm}
\title{Drug efficacy prediction in cell lines}
\date{\today}
\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: \href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Mark A. van de Wiel$^{1,3}$, Sylvia Richardson$^{3}$, \\ and Gwena{\"e}l G. R. Leday$^{3}$}

@@

	\noindent\textbf{Keywords}:

	\noindent\textbf{Software available from}:

	\section{Introduction}
	- motivation: drug response prediction from different omics data types is 
	promising both for precision medicine and drug target discovery and relates to
	actual tumor response \cite[]{iorio_landscape_2016}.

	- problem: prediction of drug efficacy in cell lines is difficult (DREAM 7 
	challenge in \cite{costello_community_2014}).

	- example data: introduction (GDSC, CCLE, NCI-60)

	- feature information: prior experts, pathways, database summaries, previous 
	experiments, type of feature (somatic mutations, copy number alterations, DNA 
	methylation, and gene expression).

	- drugs information: molecular target, drug class, pubchem, 
	omnipath DB, pharmaco DB

	- drugs x feature information

	- current literature: reviews in \cite{azuaje_computational_2017} and
	\cite{ali_machine_2019}, 
	full multivariate approach by Oslo people in \cite{mai_composite_2019}, 
	TANDEM \cite[]{aben_tandem:_2016}, DeltaNet 
	by \cite{noh_inferring_2016}, patient-derived tumor cell lines 
	in \cite{gao_high-throughput_2015}, 
	multiple kernel learning approach in \cite{costello_community_2014},
	\cite{ammad-ud-din_drug_2016} combine a 
	multiple kernel learning approach with matrix factorization, using the model 
	to uncover the latent 
	relationships between drug targets and intracellular pathways, and
	structured penalized regression in \cite{zhao_structured_2019}

	- our solution: adaptive borrowing of information through empirical Bayes, 
	guided by external feature and drug information.

	\section{Methods}
	\subsection{Simultaneous equations model}\label{sec:SEM}
	Assume we have continuous efficacy measures $y_{id}$ for cell lines 
	$i=1,\dots, n$, centred per drug $d$. We predict efficacy from molecular 
	features 
	$x_{ij}$, $j=1,\dots, p$, collected in $\x_i = \begin{bmatrix} x_{i1} & 
	\dots & x_{ip} \end{bmatrix} \tr$. We have an \textit{a priori} (possibly 
	informative) partitioning 
	of the omics features into $g=1, \dots, G$ non-overlapping groups.
	For convenience, we let 
	$g(j)$ denote the group index of molecular feature $j$. Furthermore, we have 
	characteristics of the drugs available in the form of external variables 
	$\mathbf{C} = \begin{bmatrix} \mathbf{c}_1 & \dots & \mathbf{c}_D 
	\end{bmatrix} \tr$.

	We regress the drug efficacies on the molecular feature effects $\beta_{jd}$:
	\begin{subequations}
		\begin{align}
		  y_{id} & = \x_i \tr \bbeta_d + \epsilon_{id}, \label{eq:linearmodel}\\
		  \text{with } & \epsilon_{id} \sim \mathcal{N}(0, \sigma_d^2),
		\end{align}
	\end{subequations}
	where $p$-dimensional $\bbeta_d$ are the drug-specific omics feature effects. 
	Note that (\ref{eq:linearmodel}) gives rise to a system of $d=1, \dots, D$ 
	linear simultaneous equations.

	\subsection{Bayesian prior model}
	We capture the uncertainty in the parameters through a Bayesian prior model
	and endow the parameters with the following prior:
	\begin{subequations}\label{eq:prior}
		\begin{align}
		  \beta_{jd} & \sim \mathcal{N}_p (0, \phi^2_{g(j)} \gamma_{jd}^2 
		  \sigma_d^2), 
		  \label{eq:betaprior}\\
		  \gamma_{jd}^{2} & \sim \mathcal{IG}( 
		  [(\mathbf{c}_d^j) \tr \bm{\alpha}]^{-1}, \lambda_d), \\
		  \sigma_d^{2} & \sim 1/\sigma_d^{3},
		\end{align}
	\end{subequations}
	where $\mathcal{IG}(\theta, \lambda)$ denotes an inverse Gaussian distribution 
	with mean $\theta$ and shape $\lambda$. The different molecular feature types
	are given their own weights through the $\phi^2_{g}$, while the external 
	outcome-feature data comes in through 
	our mean model for the $\gamma_{jd}^2$ hyperprior model: 
	$[(\mathbf{c}_d^j) \tr \bm{\alpha}]^{-1}$. Both the $\phi^2_{g}$ and
	$\bm{\alpha}$ are estimated in a data-driven manner. More details on the model
	itself, as well as its estimation, may be found in 
	the SM and in [! ref to first paper here].

	\section{Data}

	\subsection{GDSC}
	The GDSC project's \cite[]{yang_genomics_2013} aim is "to improve cancer 
	treatments by discovering therapeutic biomarkers that can be used to identify 
	patients most likely to respond to anticancer drug". Part of the project is to
	screen $>1000$ human cancer cell lines with drugs. The cell lines have been
	genetically characterised and several drug sensitivity measures are recorded.
	The data is freely available from \cite{garnett_systematic_2012} and consist 
	of: (i) the sensitivity measures 
	(mainly half maximal inhibitory concentration, i.e., IC50) of the cell lines 
	to the drugs, (ii) annotation of the 
	screened compounds, and (iii) the cell lines' genomic profile (mutations, 
	copy numbers, methylation profiles, and expression). We will attempt to 
	predict drug sensitivity of the cell lines, as quantified by IC50, using the
	genomic data. Other choices of efficacy measures than IC50 are possible, 
	but a discussion on the pros and cons of these different choices is not the 
	aim of this paper.

	\subsubsection{Primary data}
	We took the logarithm of the IC50 values and took the average if there were
	multiple measurements per cell line-drug combination. In the following, if we 
	refer to IC50, it denotes the log-transformed values. After iteratively 
	removing the cell line or drug with the highest proportion 
	of missings until there are no more missing values, we end up with IC50 
	estimates for 482 cell lines on 168 drugs. For the gene expression 
	data, we fit a lasso with IC50 as response to the drugs separately and 
	adjusted the penalty parameter to select approximately 300 genes per drug.

	- From Oslo:
	1.2) Preselecting GEX features: we selected gene expression features
	explaining 50\% variations over all cell lines, since this results in not too
	features and probably remains those important features.
	1.3) Preselecting CNV: we used the same strategy as Garnett et al. (2012)
	where the GDSC data are from.
	1.4) Preselecting MUT: we deleted mutated genes with less than two cell lines.

	- check out selection method in \cite{zhao_structured_2019}.

	- how to transform the data and deal with missingness? Read GDSC papers.

	\subsubsection{External data}
	Several external drug variables are available: (i) the developmental stage
	(experimental, in clinical development, or clinically approved) and (ii)
	the action (cytotoxic or targeted). We also have a external variable 
	available that relates the drugs to the features: a binary indicator whether 
	a gene is in the target pathway of the drug or not.

  - check whether compounds are PAINS \cite[]{baell_new_2010} or not. 
  PAINS indicates whether compound is frequent false hitter.

  - clinical trials database CTRP might contain info

  - extra data on GDSC might be in PharamcoGx R-package. Also contains bunch 
  of functions.

  - check GSK open target EBI or ask Chris Wallace about it. Database that links
	targets to drugs and diseases. Might be useful for drugs co-data.

	- try non-negative matrix factorisation/independent component analysis on 
	binary fingerprints to get principal components that characterise the 
	different drugs.

	\subsection{CCLE}
	- extra data on CCLE might be in PharamcoGx R-package

	\subsection{NCI-50}

	\subsection{CMap}
	- Cmap \cite[]{subramanian_next_2017,lamb_connectivity_2006} contains 
	pharma-genomic data (similar to CCLE and GDSC).

	- Data available at: 
	https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE92742.

	- Cmap website: https://clue.io/cmap.

	\section{Results}
	\subsection{GDSC}

	\section{Discussion}
	- use copula's and marginal posteriors to find multivariate posteriors for 
	contrast testing.

	- empirical Bayes gives extra information via estimation of the 
	hyperparameters, such as informativeness of the drug information (via model 
	for the $\gamma_d$) and relatedness of the tissues (via $\bm{\Omega}$ and 
	$\nu$)

	- regression on drug covariates very intuitive.

>>>>

% set options
<<settings, include=FALSE, echo=FALSE>>=
opts_knit$set(root.dir="..", base.dir="../figs")
opts_chunk$set(fig.align='center', fig.path="../figs/", 
               echo=FALSE, cache=FALSE, message=FALSE, fig.pos='!ht')
knit_hooks$set(document=function(x) {
  sub('\\usepackage[]{color}', '\\usepackage{xcolor}', x, fixed=TRUE)})
@

@@

\documentclass[a4paper,hidelinks]{article}
\usepackage{bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,
rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,
bbm,algorithm,algpseudocode,pgfplotstable,threeparttable,booktabs,mathtools,
dsfont,parskip,xr}
\externaldocument[sm-]{supplement2}

@@

% settings
\graphicspath{{../figs/}}
\pgfplotsset{compat=1.16}
\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
\setlength{\textwidth}{15cm}
\title{Empirical Bayes estimation of the normal inverse Gaussian prior in 
simultaneous-equation models}
\date{\today}
\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: 
\href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Mark A. van de Wiel$^{1,3}$, 
Sylvia Richardson$^{3}$, \\ and Gwena{\"e}l G. R. Leday$^{3}$}

@@

	\noindent\textbf{Keywords}: ...

	\section{Introduction}
	Recently, the dimensions of data sets have increased in three directions:
	(i) samples, (ii) variables, and (iii) outcomes. Here, we will 
	concern ourselves with directions (ii) and (iii). More spefically, we will 
	focus on the regression setting
	where we have multiple outcomes and a medium to high dimensional feature 
	space, i.e., $p \gtrsim n$. Standard methods
	underperform or even break down in such medium to high dimensional settings, 
	so new methodology is required.

	As a motivating example, we present an expression quantitative loci (eQTLs)
	detection problem (see Section \ref{sec:application}). In eQTLs detection, 
	the aim is to find a set of single nucleotide loci (SNPs) that explains a
	(part of a) gene's expression. The resulting heritability of the gene 
	expression traits is thought to be indicative of the relationship between 
	genotype and phenotype. eQTLs detection is often done for a set of genes 
	simultaneously and as such comprises a simultaneous-equations model (SEM) with
	multiple outcomes and a high dimensional feature set (the SNPs). Apart from
	eQTLs, SEM models are also applied in network reconstruction 
	\cite[]{kpogbezan_incorporating_2019, leday_gene_2017} 
	and drug target discovery \cite[]{noh_inferring_2016}, [! ref to second 
	manuscript].

	In many multi-outcome regression settings, extra information on both the 
	outcomes and 
	features is available. In our motivating eQTLs detection example, extra 
	information may come in the form of (i) gene annotation (e.g., length of the
	gene), (ii) SNP annotation (e.g., whether the SNP is in the gene or not), 
	(iii) SNP-gene annotation (e.g., distance of SNP to gene). The presented 
	method allows for the incorporation of such information to aid in the 
	estimation, by pooling information both across equations (genes) and 
	features (SNPs).

	We are not the first to propose to learn from external data 
	across equations and features. \cite{kpogbezan_incorporating_2019} uses 
	the presence or absence of edges in a normal tissue apoptosis pathway, to aid 
	in the estimation of a cancer tissue apoptosis pathway. Furhtermore, in an 
	eQTLs detection problem they include the location of the SNPs (inside
	the gene or not) to aid in the estimation. \cite{oman_minimax_2002} uses 
	empirical Bayes to learn from external data concerning both the outcomes and 
	the features in a low dimensional chemometrics dataset. Our proposed method,
	termed \texttt{semnig},
	extends \cite{kpogbezan_incorporating_2019,kpogbezan_empirical_2017} by 
	allowing for the inclusion of continuous external information and several 
	sources of information.

	The rest of the paper is structured as follows. In Section 
	\ref{sec:model} we introduce our model, the estimation of which is detailed
	in Section \ref{sec:estimation}. In Sections \ref{sec:simulations} and
	\ref{sec:application} we demonstrate the method in simulations and an 
  eQTLs detection problem, respectively. We end with a discussion in Section 
  \ref{sec:discussion}.

	\section{Model}\label{sec:model}
	\subsection{Simultaneous equations model}\label{sec:SEM}
	We have continuous measurements $y_{id}$ for observation $i=1,\dots, n$ on 
	outcome $d=1,\dots,D$. Throughout this paper we assume the $y_{id}$ to be 
	centred per outcome and let $\y_d = \begin{bmatrix} y_{1d} & \dots & y_{nd} 
	\end{bmatrix} \tr$. We predict the $y_{id}$ with features $x_{ij}$, 
	$j=1,\dots, p$, collected in $\x_i = \begin{bmatrix} x_{i1} & \dots & x_{ip} 
	\end{bmatrix} \tr$ and centered per feature. We have $|\bm{\alpha}|$ 
	external variables available that describes the outcome $d$, feature $j$, or
	relationship between feature $j$ and outcome $d$, collected in 
  $\mathbf{C} = \begin{bmatrix} \mathbf{c}_{1}^1 & \cdots &
	\mathbf{c}_{1}^p & \cdots & \mathbf{c}_{D}^1 & \cdots & \mathbf{c}_{D}^p
  \end{bmatrix}_{|\bm{\alpha}| \times pD} \tr$. An example of such an external
  variable from eQTL detection is the distance between SNP and gene.

	We model the outcomes $y_{id}$ as a linear function of the features $x_{ij}$:
	\begin{subequations}\label{eq:linearmodel}
		\begin{align}
		y_{id} & = \beta_{0d} + \x_i \tr \bbeta_d + \epsilon_{id},\\
		\text{with } & \epsilon_{id} \sim \mathcal{N}(0, \sigma_d^2),
		\end{align}
	\end{subequations}
	where the $p$-dimensional $\bbeta_d$ are the outcome-specific feature effects.
	Note that (\ref{eq:linearmodel}) gives rise to a system of 
	$D$ linear simultaneous equations.

	\subsection{Bayesian prior model}
	We capture the uncertainty in the parameters through a Bayesian prior model. 
	We assign a flat prior to the intercept terms $\beta_{0d}$ and integrate 
	them out. The remaining parameters are endowed with priors such that we end
	up with the following model:
	\begin{subequations}\label{eq:priormodel}
    \begin{align}
      y_{id} | \bm{\beta}_d, \sigma_d^2 & \sim \mathcal{N} 
      \(\x_i \tr \bm{\beta}_d, \sigma_d^2\), \\
      \beta_{jd} | \gamma_{jd}^2, \sigma_d^2 & \sim \mathcal{N} 
      \(0, \sigma_d^2 \gamma_{jd}^2 \), \label{eq:betaprior} \\
      \gamma_{jd}^2 & \sim \mathcal{IG}
      \( [(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^{-1}, \lambda_{d} \), \\
      \sigma_d^2 & \sim 1/\sigma_d^3,
    \end{align}
  \end{subequations}
  where $\mathcal{IG} ([(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^{-1}, \lambda_{d})$ 
  denotes the inverse Gaussian distribution with mean 
  $[(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^{-1}$ and shape $\lambda_d$.

  The normal inverse Gaussian (NIG) prior model for $\beta_{jd}$ and 
  $\gamma^2_{jd}$ in (\ref{eq:priormodel}) was introduced in 
  \cite{barndorff-nielsen_hyperbolic_1978} and is, since 
  \cite{barndorff-nielsen_normal_1997}, routinely applied in mathematical 
  finance (see, e.g., \cite{kalemanova_normal_2007}). Supplementary Material
  (SM) Section \ref{sm-sec:nigprior} contains more details on this prior.
  In Figure \ref{fig:dens_igaussian_marginalbeta}, the NIG 
  prior distribution, marginalised over $\gamma_{jd}^2$, is depicted together 
  with three other common choices,
  the Student's $t$ (see Section \ref{sec:inversegammamodel} for more details on 
  this prior), normal (ridge) and Laplace (lasso) priors, 
  all scaled to variance one. 
  From this Figure we see that, depending on the choice
  of hyperparameters, the NIG prior can vary from heavier tails than the lasso, 
  to more Gaussian tails like the ridge, but behaves similarly to the 
  Student's $t$ prior. Our argumentation to model the 
  $\gamma^2_{jd}$ by an inverse Gaussian 
  distribution, as has been suggested in \cite{fabrizi_specification_2016} and 
  \cite{caron_sparse_2008}, is two-fold: (i) the NIG model is more flexible than
  the standard ridge and lasso priors (as seen from Figure 
  \ref{fig:dens_igaussian_marginalbeta}), and (ii) the NIG prior allows to model
  the mean as a 
  function of the external variables $\mathbf{c}_{d}^j$ more conveniently than 
  the Student's $t$ prior, as explained in Section 
  \ref{sec:empiricalbayes}. 
<<dens_igaussian_marginalbeta, fig.cap="Several choices of marginal $\\beta_j$ prior, scaled to $\\V(\\beta_j)=1$.", out.width="60%", fig.asp=3/4>>=
@

  \cite{moran_variance_2018} argue that in regression models, $\bm{\beta}_d$ 
  and $\sigma_d^2$ should be independent \textit{a priori}, i.e.
  $\beta_{jd} | \gamma_{jd}^2, \sigma_d^2 \sim \beta_{jd} | 
  \gamma_{jd}^2$. However, in our case
  that is debatable. The $D$ regressions need not be on the same scale, so in
  order to jointly consider the $\bm{\beta}_d$ some 
  form of calibration is required. A straightforward way 
  of doing this is to scale the $\bm{\beta}_d$ prior with 
  $\sigma_d^2$ as in (\ref{eq:betaprior}), resulting in a 
  scale-free prior with respect to the response. Apart from the scaling,
  in this setting, the conjugate prior model is preferred
  over the non-conjugate model, because the variational Bayes approximation 
  (see Section \ref{sec:estimation}) assumes independent $\sigma_d^2$ and 
  $\bm{\beta}_d$ in the posterior, which is actually true for 
  the conjugate model. Therefore, the approximate expectation used in the 
  empirical Bayes step (see Section \ref{sec:estimation}) is more accurate.

  A few remarks on the choice of error variance prior are justified here: 
  many authors endow error variance components with vague gamma priors. 
  \cite{gelman_prior_2006}, 
  among others, advises against this practice. The degree of `vagueness' has a 
  large influence on the posterior, while degree of `vagueness' is a difficult 
  parameter to set. This influence is especially pronounced if the likelihood is
  relatively flat, as may be reasonably expected in the large $p$, small $n$ 
  setting we are in. We therefore model the error variance with Jeffrey's 
  objective prior 
  \cite[]{jeffreys_invariant_1946}. In the derivation of our Jeffrey's prior
  for the error variance, we jointly consider an unknown mean and variance. This
  joint consideration results in the somewhat unorthodox $1/\sigma^3$ Jeffrey's
  prior.

  Several extensions of the model are described in [! ref to second paper here].

	\section{Estimation}\label{sec:estimation}
	\subsection{Variational Bayes}\label{sec:variationalbayes}
	The posterior corresponding to the model described in (\ref{eq:linearmodel})
	and (\ref{eq:priormodel}) is not available in closed form. We therefore 
	approximate the posterior by variational Bayes, where we force the approximate
	posterior 
	density to factorise as: $p(\bbeta_d, \sigma_d^2, \gamma_{1d}^2, \dots, 
	\gamma_{pd}^2) \approx 
	Q_d(\cdot) = q(\bbeta_d) \cdot q(\sigma_d^2) \cdot \prod_{j=1}^p 
	q(\gamma_{jd}^2)$. For
	notational convenience, we slightly abuse notation and let $q(\cdot)$ denote
	different densities for different inputs.
	Under such a factorisation, the marginal variational posteriors that 
	minimise the Kullback-Leibler divergence of the true posterior to the 
	variational Bayes approximation \cite[]{neal_view_1998} are given by:
	\begin{align*}
    q(\bm{\beta}_d) & \overset{D}{=} \mathcal{N}_p 
    (\bm{\mu}_d, \bm{\Sigma}_d), \\
    q(\gamma_{jd}^2) & \overset{D}{=} \mathcal{GIG}
    \(-1, \lambda_{d}[(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^2, \delta_{jd}\), \\
    q(\sigma^2_d) & \overset{D}{=} \Gamma^{-1} \(\frac{n + p + 1}{2}, \zeta_d\).
  \end{align*}
	See SM Section \ref{sm-sec:vbderivations} for the derivations. The variational
	parameters $\bm{\mu}_d$, $\bm{\Sigma}_d$, $\delta_d$, and
	$\zeta_d$ contain cyclic dependencies and are iteratively 
	updated by:
  \begin{align*}
    \bm{\Sigma}_d^{(h+1)} & = (a_d^{(h)})^{-1} 
    \[\X \tr \X + \diag(b_{jd}^{(h)})\]^{-1}, \\
    \bm{\mu}_d^{(h+1)} & = \[\X \tr \X + \diag(b_{jd}^{(h)})\]^{-1} 
    \X \tr \y_d, \\
    \delta_{jd}^{(h+1)} & = a_d^{(h)} \[(\bmu^{(h+1)}_{jd})^2 + 
    (\bSigma_d)_{jj}\] + \lambda_{d}, \\ 
    \zeta_d^{(h+1)} & = \frac{1}{2} \bigg[\mathbf{y}_d \tr \mathbf{y}_d -
    2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} + 
    \trace ( \X \tr \X \bm{\Sigma}_d^{(h+1)}) + 
    (\bm{\mu}_d^{(h+1)}) \tr \X \tr \X \bm{\mu}_d^{(h+1)} \\
    & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + 
    \trace \[ \diag (b_{jd}^{(h+1)}) \bm{\Sigma}_d^{(h+1)}\] + 
    (\bm{\mu}_d^{(h+1)}) \tr \diag (b_{jd}^{(h+1)}) 
    \bm{\mu}_d^{(h+1)}\bigg],
  \end{align*}
  until convergence. Here, we set
  \begin{align}
    a_d^{(h)} & =\E_{Q^{(h)}}(\sigma_d^{-2})=(n + p + 1)/(2 \zeta_d^{(h)}), 
    \nonumber\\
    b_{jd}^{(h)} & = \E_{Q^{(h)}}(\gamma_{jd}^{-2}) = 
    \sqrt{\frac{\lambda_{d}[(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^2}
    {\delta_{jd}^{(h)}}}
    \frac{K_0(\sqrt{\delta_{jd}^{(h)}\lambda_{d}}
    [(\mathbf{c}_{d}^j) \tr \bm{\alpha}])}
    {K_1(\sqrt{\delta_{jd}^{(h)}\lambda_{d}}
    [(\mathbf{c}_{d}^j) \tr \bm{\alpha}])} + 
    \frac{2}{\delta_{jd}^{(h)}}, \label{eq:ratiomodifiedbessel}
  \end{align}
  where $K_{\nu}(x)$ denotes the modified Bessel function of the second kind.
  A method for fast and numerically stable calculation of ratios of
  modified Bessel functions of the second kind, as in 
  (\ref{eq:ratiomodifiedbessel}), is given in SM Section 
  \ref{sm-sec:ratiosmodifiedbessels}.

  \subsection{Empirical Bayes}\label{sec:empiricalbayes}
  To make use of the external data, we parametrise the prior mean of the
  $\gamma_{jd}^2$ as $[(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^{-1}$. 
  This allows us to
  include continuous external variables $\mathbf{c}_{d}^j$ into the model. 
  Additionally, it reduces the number of hyperparameters from $pD$ to 
  $|\bm{\alpha}| + D$. A full Bayesian model
  then requires the specification of the hyperparameters 
  $\bm{\alpha}$ and $\blambda = \begin{bmatrix} \lambda_1 & \cdots & \lambda_D 
  \end{bmatrix} \tr$. These are abstract and hard to interpret parameters 
  for which we generally lack expert knowledge. They do, however, have a 
  significant influence on the shape of the posterior distribution. We 
  therefore propose to estimate these hyperparameters by empirical Bayes. 
  Simply put, empirical Bayes fits the prior model to the data and is, as such, 
  objective Bayes (up to model specification). In our case, this results in an   
  objective and data-driven inclusion of the external data.

	The canonical method for empirical Bayes is maximisation of the marginal 
	likelihood with respect to the hyper parameters. In 
	\cite{casella_empirical_2001} the marginal likelihood is maximised by an EM 
	algorithm:
	\begin{align*}
	  \bm{\alpha}^{(l+1)},\bm{\lambda}^{(l+1)} & = \underset{\bm{\alpha},
	  \bm{\lambda}}{\argmax}\E_{\cdot | \Y} 
	  [\log p(\Y, \B, \bm{\Gamma}^2, \bsigma^2) | \bm{\alpha}^{(l)},
	  \bm{\lambda}^{(l)}] \\
	  & = \underset{\bm{\alpha},\bm{\lambda}}{\argmax} 
	  \E_{\cdot | \Y} [\log \pi (\bm{\Gamma}^2) | \bm{\alpha}^{(l)},
	  \bm{\lambda}^{(l)}],
	\end{align*}
	where $\Y = \begin{bmatrix} \y_1 & \cdots & \y_D \end{bmatrix}$,
	$\B = \begin{bmatrix} \bbeta_1 & \cdots & \bbeta_D \end{bmatrix}$, 
	$\bsigma^2 = \begin{bmatrix} \sigma^2_1 & \cdots & 
	\sigma^2_D \end{bmatrix} \tr$,
	and $\bm{\Gamma}^2 = \begin{bmatrix} \bgamma^2_1 & \cdots & \bgamma^2_D 
	\end{bmatrix}\tr$, with 
	$\bgamma^2_d = \begin{bmatrix} \gamma^2_{1d} & \cdots & \gamma^2_{pd} 
	\end{bmatrix}\tr$,
	and the expectation is with respect to the joint posterior. In our case, this 
	posterior is not available in closed form, which renders the expectation 
	difficult. While \cite{casella_empirical_2001} suggests to approximate the 
	expectation by a Monte Carlo sample, we propose to use the variational Bayes 
	approximation developed in Section \ref{sec:variationalbayes}:
	$$
	\bm{\alpha}^{(l+1)},\bm{\lambda}^{(l+1)} = 
	\underset{\bm{\alpha},\bm{\lambda}}{\argmax}\E_{Q^{(l)}} 
	[\log \pi(\bm{\Gamma}^2)],
	$$
	where now the expectation is with respect to the converged variational 
	posterior $Q^{(l)}=\prod_{d=1}^D Q_d^{(l)}$. The resulting estimating 
	equations are:
  \begin{align*}
    \bm{\alpha} & = \[ \mathbf{C} \tr \bLambda \diag 
    (e_{jd}^{(l)}) \mathbf{C} \]^{-1} \mathbf{C} \tr \bLambda
    \mathbf{1}_{pD \times 1}, \\
    \lambda_{d} & = p \left\{ \sum_{j=1}^p b_{jd}^{(l)} + 
    \sum_{j=1}^p e_{jd}^{(l)} [(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^2 - 
    2 \sum_{j=1}^p (\mathbf{c}_{d}^j) \tr \bm{\alpha} \right\}^{-1},
  \end{align*}
  where $\bLambda=\diag(\lambda_d) \otimes \I_p$
  and $e_{jd}^{(l)} = \E_{Q^{(l)}}(\gamma_{jd}^2) = (b_{jd}^{(l)} - 
  2/\delta_{jd}^{(l)}) \cdot \delta_{jd}^{(l)}/[\lambda_{d}^{(l)}
  \left\{(\mathbf{c}_{d}^j) \tr \bm{\alpha}^{(l)}]^2\right\}$. 
  Solving the equations is done by 
  iteratively reweighted least squares.

  Variaional Bayes approximations are known to underestimate posterior 
  variances \cite[]{rue_approximate_2009,consonni_mean-field_2007,
  bishop_pattern_2006,wang_inadequacy_2005}. If posterior variances are 
  required, 
  we suggest generating samples from the posterior after the empirical Bayes 
  iterations have converged. A Gibbs sampler to do so is described in 
  the SM Section \ref{sm-sec:gibbssampler}. Alternatively, the \texttt{R} 
  package \texttt{rstan} \cite[]{guo_rstan:_2018} allows the user to sample
  from a generic posterior. An implementation of model
  (\ref{eq:priormodel}) in \texttt{stan} code may be found on 
  \url{https://github.com/magnusmunch/cambridge}.

  Note that so far, the design matrix $\X$ is assumed to be shared across the
  equations. This is not a requirement. That is, we may have D different design 
  matrices $\X^d$ that need not have the same number of 
  columns, i.e., we have $p_d$ instead of $p$. This becomes useful in, e.g. 
  eQTLs detection, where we regress genes on a gene-specific selection
  of SNPs, often based on SNP-gene distance.

  \section{Simulations}\label{sec:simulations}
  We conducted a simulation study consisting of three parts investigating (i) 
  the NIG prior performance, (ii) the variational Bayes approximation, and (iii)
  the empirical Bayes estimation.

  \subsection{NIG prior performance}
  \label{sec:inversegammamodel}
  The aim of this part is to investigate the 
  performance of the NIG prior model. Since we may do so in the univariate 
  setting, so we drop all $d$ indices in the following. 
  A common default prior for latent variance components is the inverse Gamma 
  prior:
  $$
  \gamma_j^2 \sim \Gamma^{-1}(\eta/2,\xi/2),
  $$
  with shape $\eta/2$ and scale $\xi/2$, both rescaled to match our 
  inverse Gaussian prior formulation. The resulting marginal prior for the model
  parameters
  (conditional on the error variance $\sigma^2$) is then a scaled Student's 
  $t$-distribution:
  $$
  \beta_j | \sigma^2 \sim \text{Student's } 
  t\(\eta,\sigma \sqrt{\xi/\eta}\),
  $$
  with degrees of freedom $\eta$ and scale $\sigma \sqrt{\xi/\eta}$.
  See SM Section \ref{sm-sec:inversegamma} for more details on this prior. 
  We compare the 
  posteriors of the proposed NIG to the scaled Student's 
  $t$-distribution prior in a simulated setting.

  We compare the posterior distributions to the true posterior for a 
  number of measures. Divergence between the distributions
  is quantified through a multivariate extension of the Cram{\'e}r-von Mises
  criterion \cite[]{baringhaus_new_2004}, where smaller values of the criterion
  indicate a better fit. We also assess 95\% coverage of the 
  posterior credible intervals. The accuracy of the 
  point estimates of $\beta_j$ in the form of posterior means is assessed by
  the mean squared error (MSE) and correlation between the point estimates 
  and the true $\beta_j$.

  We studied two settings: one dense and one sparse. In the dense setting, 
  we simulated $n=100$ features $\x_i$ from a $p=200$ dimensional Gaussian 
  distribution with zero mean vector, variances equal to one, and all 
  covariances set to $\rho=0.5$. The outcomes were generated through $y_i = 
  \x_i\tr \bbeta + \epsilon_i$, with the $\epsilon_i$ from the standard normal 
  distribution. 
  Likewise, we generated the model parameters $\beta_j$ from a standard normal
  distribution.

  We choose the hyperparameters such that we get all combinations of marginal
  prior variances and kurtoses $\V(\beta_{j} | \sigma^2) \in 
  \{ 1/2, 1, 2 \}$ and $\mathcal{K}(\beta_{j} | \sigma^2) \in 
  \{ 4, 10 \}$. This leads to the settings in Table \ref{tab:hyper_set1}. For 
  all priors, we generated 1000 draws (after 1000 burnin) from the 
  posteriors using an MCMC sampler implemented in \texttt{rstan}. We repeated
  the simulation 100 times and report 
  results in Figures \ref{fig:boxplots_igaussian_res4.1_post1} and 
  \ref{fig:boxplots_igaussian_res4.1_post2}. 
<<hyper_set1>>= 
set <- read.table("results/simulations_igaussian_set4.csv")
kableExtra::add_header_above(kableExtra::kable_styling(
  knitr::kable(set[, c("ctalphainv", "lambda", "eta", "xi", "vars", "kurts")],
               col.names=c("$(\\mathbf{c} \\tr \\bm{\\alpha})^{-1}$",
                           "$\\lambda$", "$\\eta$", "$\\xi$",
                           "$\\V(\\beta_{j} | \\sigma^2)$",
                           "$\\mathcal{K}(\\beta_{j} | \\sigma^2)$"),
               caption="Hyperparameter settings",
               "latex", booktabs=TRUE, escape=FALSE, digits=2),
  latex_options=c("HOLD_position")),
  header=c(" "=1, "NIG"=2, "Student's t"=2, "Prior"=2))
@

<<boxplots_igaussian_res4.1_post1, fig.cap="Comparison of the NIG and Student's t priors in the dense setting in (a)-(d) simulation setting 1, (e)-(h) setting 2, and (i)-(l) setting 3, through Cramér-von Mises criterion between the model and true posteriors, correlation and MSD between the posterior means and true parameters, and 95\\% credible interval coverage of the true values.", out.width="100%", fig.asp=1>>=
@

<<boxplots_igaussian_res4.1_post2, fig.cap="Comparison of the NIG and Student's t priors in the dense setting in (a)-(d) simulation setting 4, (e)-(h) setting 5, and (i)-(l) setting 6, through Cramér-von Mises criterion between the model and true posteriors, correlation and MSD between the posterior means and true parameters, and 95\\% credible interval coverage of the true values.", out.width="100%", fig.asp=1>>=
@

  [! describe results here]
